Thermodynamic response functions in a cell fluid model

### 2024-09-03 main_morse_response_functions.tex
add explicit derivative calculation for beta_V

	At $T>T_c$ the equation of state of a cell fluid model in terms of the density and the temperature has the following form
	\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\eta (\bar n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4,
	\end{equation}
 The function $E_\nu (\bar n, T)$ in \eqref{eq:ro_nT} is the function $E_\mu (M,T)$ \eqref{eq:E_mu} rewritten in terms of density and temperature taking into account the expression \eqref{eq:M_nT}. For the reduced pressure we write
 \begin{equation}
 	\label{eq:eosPTn_reduced}
 	P^* = (1 + \tau)T^*_c \left(E_\eta (\bar{n},T) + \bar{M} \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4\right).
 \end{equation}

@@

	We introduce the reduced thermal pressure coefficient by
	\begin{equation}
		\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\beta^*_V = \frac{1}{T^*_c} 
		\left[ \left(\frac{\partial P^*}{\partial \tau}\right)_{\mu} 
		- \left(\frac{\partial P^*}{\partial \mu}\right)_T 
		\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T} 
		\right].
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the thermal pressure coefficient follow immediately from the definition~\eqref{def:therm_pres_coef}
	\begin{equation}
		\beta_V = \left(\frac{\partial P}{\partial T}\right)_\rho
	\end{equation}
	and
	\begin{equation}
		\beta^*_V = \frac{1}{T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho}
	\end{equation}

@@

	\section{\label{sec:app-a} Explicit expressions for derivatives}
	In this appendinx, explicit expressions are presented for derivatives of quantities entering the equation of state, with respect to temperature. This expressions are useful while calculating the response functions in Section~\ref{sec:respons_functions}.

	For the quantity $\bar{M}$ from Eq.~\eqref{eq:M_nT} one has
	\begin{equation}
		\left( \frac{\partial \bar{M}}{\partial \tau} \right)_{\bar{n}} =
		\frac{k_{\rm B}T_c \rho_n}{W(0)} \frac{1 - \omega_0}{1 + \tau\omega_0},
	\end{equation}
	where
	\begin{equation*}
		\omega_0 = \frac{\chi_0 + A_{\gamma}}{B - 1 + \chi_0}.
	\end{equation*}
	For $\tilde{D}(0)$ from~\eqref{def:D0} one has
	\begin{equation}
		\frac{\partial \tilde{D}(0)}{\partial \tau} = -\frac{k_{\rm B}T_c}{W(0)} \frac{1 - \omega_0}{1 + \tau \omega_0}.
	\end{equation}

>>>>

	At $T>T_c$ the equation of state of a cell fluid model in terms of the density and the temperature has the following form
	\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\eta (\bar n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4,
	\end{equation}
 The quantity $E_\nu (\bar n, T)$ in \eqref{eq:ro_nT} is the function $E_\mu (M,T)$ \eqref{eq:E_mu} rewritten in terms of density and temperature taking into account the expression \eqref{eq:M_nT}
	\begin{eqnarray}\label{eq:E_nu}
 		E_\nu (\bar{n}, T) & = & - \frac{\ln (2\pi \beta W(0))}{2 N_v}  +  g_0 
 		\nonumber\\
 		&& - \frac{\beta W(0)}{2} 
 		\left(\bar{M} - g_1 + \frac{g_3}{g_4} \tilde{D}(0) + \frac{g_3^3}{6g_4^2} \right)^{2} 
 		\nonumber\\
 		&& - \frac{g_3}{g_4} {\bar{M}} - \frac{g_3^2}{2 g_4^2}  \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
 	\end{eqnarray}
 For the reduced pressure we write
 \begin{equation}
 	\label{eq:eosPTn_reduced}
 	P^* = (1 + \tau)T^*_c \left(E_\eta (\bar{n},T) + \bar{M} \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4\right).
 \end{equation}

@@

	We introduce the reduced thermal pressure coefficient by
	\begin{equation}
		\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\label{eq:beta_star_m}
		\beta^*_V = \frac{1}{T^*_c} 
		\left[ \left(\frac{\partial P^*}{\partial \tau}\right)_{\mu} 
		- \left(\frac{\partial P^*}{\partial \mu}\right)_T 
		\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T} 
		\right].
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the thermal pressure coefficient follow immediately from the definition~\eqref{def:therm_pres_coef}
	\begin{equation}
		\beta_V = \left(\frac{\partial P}{\partial T}\right)_\rho
	\end{equation}
	and
	\begin{equation}
		\label{eq:beta_star_n}
		\beta^*_V = \frac{1}{T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho}
	\end{equation}

@@

	\section{\label{sec:app-a} Explicit expressions for derivatives}
	In this appendix, explicit expressions are presented for derivatives of quantities entering the equation of state, with respect to temperature. This expressions are useful while calculating the response functions in Section~\ref{sec:respons_functions}.

	For the quantity $\bar{M}$ from Eq.~\eqref{eq:M_nT} one has
	\begin{equation}
		\left( \frac{\partial \bar{M}}{\partial \tau} \right)_{\bar{n}} =
		\frac{k_{\rm B}T_c \rho_n}{W(0)} \frac{1 - \omega_0}{1 + \tau\omega_0},
	\end{equation}
	where
	\begin{equation*}
		\omega_0 = \frac{\chi_0 + A_{\gamma}}{B - 1 + \chi_0}.
	\end{equation*}
	For $\tilde{D}(0)$ from~\eqref{def:D0} one has
	\begin{equation}
		\frac{\partial \tilde{D}(0)}{\partial \tau} = -\frac{k_{\rm B}T_c}{W(0)} \frac{1 - \omega_0}{1 + \tau \omega_0}.
	\end{equation}
	While calculating both derivatives, we take into account
	\begin{equation}
		\frac{\partial}{\partial \tau} \left[\beta W(0)\right] = -\frac{\beta_c \Phi^{(r)}(0)}{(1+\tau)^2} (B - 1 - A_{\gamma}).
	\end{equation}
	The derivative of the quantity $E_\nu$ from~\eqref{eq:E_nu} with respect to temperature $\tau$ is
	\begin{eqnarray}
		\frac{\partial E_\nu}{\partial \tau} & = &  -\frac{1}{2}\left(\bar{M} - g_1 + \frac{g_3}{g_4} \tilde{D}(0) + \frac{g_3^3}{6g_4^2} \right)^{2} 
		\frac{\partial \beta W(0)}{\partial \tau} 
		\nonumber\\
		&&-\beta W(0)\left(\bar{M} - g_1 + \frac{g_3}{g_4} \tilde{D}(0) + \frac{g_3^3}{6g_4^2} \right)
		\nonumber\\
		&& \times \left(\frac{\partial \bar{M}}{\partial \tau} + \frac{g_3}{g_4}\frac{\partial\tilde{D}(0)}{\partial \tau}\right)
		\nonumber\\
		&& -\frac{g_3}{g_4}\frac{\partial \bar{M}}{\partial \tau} 
		- \frac{g_3^2}{2g_4^2}\frac{\partial \tilde{D}(0)}{\partial \tau}.
	\end{eqnarray}
	Collecting these all formulas together, we can explicitly calculate $\beta^*_V$ from~\eqref{eq:beta_star_n}
	\begin{eqnarray}
		\beta^*_V & = & E_\eta (\bar{n},T) + \bar{M} \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4
		\nonumber\\
		&& + (1+\tau) \left(\frac{\partial E_{\nu}}{\partial \tau} + \frac{\partial \bar{M}}{\partial \tau}\rho_{n} + \frac{\rho_n^2}{2}\frac{\partial \tilde{D}(0)}{\partial \tau}\right).
	\end{eqnarray}
